Adaptive broadcast tree for multicores

### 2013-04-29 paper.tex
Rewrote Introduction.
New version is much more concrete: automatic configuraiton of atomic broadcast for diverse and complex multicores

\title{Replication on multicore considering hierarchy and
  characteristics of the network}

@@

There are two problems:
\begin{enumerate}
\item Performance of distributed algorithms on multicores. Programmers
  don't get it right due to different complexity measures. We deal
  with this by improving these algorithms. This will lead to higher
  performance of existing algorithms.
\item Diversity of hardware. Hand-tuning algorithms is tedious. Need
  to do this automatically. To evaluate this, we compare the average
  performance of our automatic configuration on a wide range of
  multicores with performance of a single hand-tuned implementation.
\end{enumerate}

\paragraph{What do you want to enable?} Increase algorithm performance
on multicores by applying distributed principles to overcome
scalability challenges on such machines.

\paragraph{What problem are you solving, and why is it hard?} We are
applying distributed algorithms to multicores to enable better
scalability. However, characteristics of a multicore a different to
those of traditional distributed systems. Traditional algorithms need
to be reevaluated for multicores.

\paragraph{What's the related work?} People apply techniques from
distributed systems to deal with scalability challenges on multicores,
e.g.\ databases~\cite{Salomie2011, Wiesmann2000} and operating
systems~\cite{fos:osr09, tornado:osdi99, barrelfish:sosp09}. But they
don't to it properly. They do not consider machine characteristics or
apply the wrong distributed technologies. Also, what they do is rather
ad hoc.

\paragraph{What new idea(s) will solve the problems?} Instead of just
reusing distributed algorithms, we reevaluate them for multicores and
investigate the impact of the multicore network model on algorithm
performance.

\paragraph{How will you go about it?} We investigate the problem for
atomic-broadcasts since many problems in distributed systems can be
reduced to this communication primitive. Traditional implementations
are sending requests to all nodes in parallel to hide high propagation
delays. %
However, this does not work on multicores, since the propagation time
is not dominating any longer. %
Instead, we build a tree for communication to parallelize transmit and
receive cost across nodes. This has been done before (BF SOSP, IBM,
.. ), but not for the general case of multicore communication.

\paragraph{How will you know and show it works?} We will use a
microbenchmark measuring broadcast cost and compare our tree
implementation to the sequential case. Maybe we put some high-level
applications on top of it.

\paragraph{What is your hypothesis?} Performance of distributed
algorithms can be improved if characteristics of multicores are
considered when building and/or selecting them. %
For the broadcast example, this means: the time complexity for a
networks with dominating transmit and receive time will be $O(log(n))$
instead of $O(n)$.

>>>>

\title{Adaptive broadcast tree for multicores}

@@

\begin{enumerate}
\item Performance of distributed algorithms on multicores. Programmers
  don't get it right due to different complexity measures. We deal
  with this by improving these algorithms. This will lead to higher
  performance of existing algorithms.
\item Diversity of hardware. Hand-tuning algorithms is tedious. Need
  to do this automatically. To evaluate this, we compare the average
  performance of our automatic configuration on a wide range of
  multicores with performance of a single hand-tuned implementation.
\end{enumerate}
We deal with the latter one for now.

% 
Atomic broadcast is a building-block for many higher-level distributed
algorithm such as state-machine replication. Such distributed
algorithms are increasingly used to overcome scalability challenges on
multicore machines. Examples are databases~\cite{Salomie2011,
  Wiesmann2000} and operating systems~\cite{fos:osr09, tornado:osdi99,
  barrelfish:sosp09}.

% What do you want to enable?
We want to investigate how to achieve better performance of atomic
broadcast across a wide range of multicore machines without manually
tuning the implementation to concrete machines.

% What problem are you solving, and why is it hard?
The problem is complexity and diversity of modern multicore
machines. Multicores come in many flavors depending on vendor and
generation of the machine. All of these machines have fundamentally
different characteristics such as number of nodes, NUMA
characteristics, interconnect topologies, propagation time on
interconnect and use of shared caches. These characteristics are hard
to understand, but important to consider for application performance.

% Related work
\stefan{Fill in related work: in a nutshell, they do tuning, but only
  for specific machines}

% New ideas
Instead of manually tuning algorithms to characteristics of individual
machines, we automatically configure our atomic broadcast
implementation based on a machine model.
% How will you go about it?
We will base our atomic broadcast on overlay networks, which we
automatically configure based on the machine model. One questions is
which information should actually be encoded in the machine model. %
Our idea is to then explore different approaches (such as MST
algorithms, clustering) to \emph{find good overlay networks} based on
the network model to base the atomic broadcast on. Then, for every
node, we need to decide which message to send first. This is a
\emph{scheduling} problem.

% How do we show it works
We auto-configure a wide range of different multicore machines and
show the performance of our implementation comes close to the
performance of hand-tuned implementations for all of these machines.

% Hypothesis
My hypothesis is that automatic tuning of atomic broadcast (and other
distributed algorithms) to machine characteristics is worth
doing. Obviously, we cannot achieve performance of hand-tuned
implementations, but we can get close enough to argue that loosing a
bit of performance to avoid tedious manual reconfiguration is worth if
for many applications (while others, where performance \emph{really}
matters, might still want to manually configure their stuff).
